---
title: "ED Attendance report: `r site_name`"
output:
   html_document:
    toc: yes
    toc_depth: 3
    toc_float: 
     collapsed: false     
date: "2024-09-02"
theme: Cerulean
params: 
  site_name: site_name
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

site_name <- params$site_name

```

<style type="text/css">
.main-container {
  max-width: 80% !important;
  margin: auto;
}
</style>


### Abstract

This study is aimed at identifying large changes in presentations over
time across different sites, to identify where we have the most
statistically different increases. In addition a cohort of patients is
identified that could potentially benefit from non ED services.

This is an analysis of ED attendances across the SW region, focusing on
the 5 most common, non ambulance conveyances to type 1 departments. The
data has been analysed using XXX methodology to identify avoidanble
attendances. A sub group of patients who are not classified as avoidable
attendances or subsequently become admissions has been identified, these
are patients who are currently appropriate for ED but only have minor
investigations and procedures before discharge. We have analysed the
demographics of these patients for personal characteristics as well as
their attendance patterns to help identify a patient group that an
alternative provision outside of ED could be developed.

### Method

Data was taken from ECDS for the last 18 months of non ambulance
conveyances. The top 5 presentations per site were identified by number
of attendances using SNOMED codes. The dataset was filtered to remove
admitted patients. Avoidable attendances were identified useing XXX
methodology. The data was further analysed to create a cohort of
patients who were of minor acuity and recieved low level investigations
and treatments.

```{r, echo=FALSE,  fig.height=6,fig.width = 16}
flow


```

This cohort was then further analyised to identify characteristics of
these patients using cluster analysis to try to identify key groups of
patients. This included demographic, geographic and time attendance
patterns.

A resource time was created by removing the wait time from the total
time in department to give a proxy for resource utilisation time in
department.

This report utilises an avoidable attendances methodology to help
identify a middle ground of patients. These are a cohort of patients who
are non ambulance conveyances to a type 1

### Benchmarking change over time {.tabset .tabset-fade .tabset-pills}

These charts look at the last 6 months of attendances and compare them to the same 6 months a year ago.  The percentage change is then calculated.  This is broken down across the various attendance types we have identified.  

We are most interested in this study for the change in divertible attendances.The other types are provided for context.  Further in the report we can see this in a more traditional time series for this individual site.

#### Total Attendances
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_to


```

#### Avoidable Attendances
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_aa


```

#### Divertable Attendances
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_da


```


#### Appropriate Attendances (Discharged)
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_ap 


```


#### Appropriate Attendances (Admissions)
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_ad 


```



### Appendix:

Data sources: [Reporting].[MESH_ECDS_EC_Core]
,[Reporting_UKHD_ODS].[Provider_Site], [AGEM_TravelTime].[LSOA_to_Site]
Time period: 18 months from current date Avoidable attendances is based
on work by NHS Digital in 2020:
<https://web.archive.org/web/20210226213236/https:/digital.nhs.uk/data-and-information/data-tools-and-services/data-services/innovative-uses-of-data/demand-on-healthcare/unnecessary-a-and-e-attendances>
Resource time: Total time to discharge - Time to treat. 95th percentile
was used to account for outliers. Any patients where their time to treat
was longer than their discharge time were removed.

```{r pressure, echo=FALSE}
data_summary_table
```

### Benchmarking

Top 5 presenations for this site

Tab tab tab

```         
                  Ateendances                               Resource Time
          18ths  12 mths 6 mths 0mths    Change
```

This Site Total Avoidable Alt cohort Other cohort Admissions

### Resource time {.tabset .tabset-fade .tabset-pills}

#### Resource time site comparison

The below chart shows the average resource time per patient for each
site over the last 18 months.

Guide for interpreting the box plot:

**Key Components of a Box Plot:**

**Box:** The central rectangle represents the interquartile range (IQR),
which contains the middle 50% of the data.

**Median:** The line within the box indicates the median value,
separating the lower and upper halves of the data.

**Whiskers:** The lines extending from the box are called whiskers. They
typically reach to the minimum and maximum values excluding outliers.

**Outliers:** Data points that fall outside of the whiskers, represented
by the dots, are considered outliers.

```{r echo=FALSE, results='hide',fig.asp = 0.8, fig.height=10,fig.width = 12,message=FALSE, warnings = FALSE}
##box plot by year

```

#### Resource time site level time comparison

The below chart shows a comparison of three six month time periods.

```{r echo=FALSE, results='hide',fig.asp = 0.8, fig.height=10,fig.width = 12,message=FALSE, warnings = FALSE}



```

#### Resource time SPC

The below SPC chart shows the median resource time per patient over the
latest 18 month for this site.

```{r echo=FALSE, results='hide',fig.asp = 0.8,  fig.height=10,fig.width = 12,message=FALSE, warnings = FALSE}


```

### Chief Complaint

The below SPC chart shows the proportion of attendances by each
complaint time over the latest 18 month for this site.

```{r echo=FALSE, results='hide',fig.asp = 0.8, fig.height=10, fig.width = 12,message=FALSE, warnings = FALSE}


```

