---
title: "ED Attendance report: `r str_to_title(site_name)`"
output:
   html_document:
    toc: yes
    toc_depth: 3
    toc_float: 
     collapsed: false     
date: "2024-09-02"
theme: Cerulean
params: 
  site_name: site_name
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

site_name <- params$site_name

```

<style type="text/css">
.main-container {
  max-width: 80% !important;
  margin: auto;
}
</style>


## Abstract

This report identifies changes in presentations over time at this site for emergency department attendances who were not conveyed by ambulance. Patients were categorised into one of four cohorts; the report focuses primarily on the cohort who could potentially be diverted from emergency department care if alternative services were available. 

The patients are split into those categorised as **avoidable attendances**, **divertible attendances**, **appropriate attendances** and **admissions**. 

The report benchmarks the last 6 months to the same  6 months a year ago (so that we do not compare summer to winter), and explores the increases or decreases by presentation for each of the above categories.

A further breakdown and cluster analysis focuses on the **divertible attendances** to see if there is any insight within the demographic, temporal and geographic characteristics of those patients that may support developing alternative pathways or identification of these patients.

## Method

Data was taken from ECDS for the last 18 months of non ambulance conveyances. The top 6 presentations per site were identified by number of attendances using chief complaint based on SNOMED codes. 

The attendances were then classified as follows:


```{r, echo=FALSE,  fig.height=6,fig.width = 16}
flow
```


-  **Avoidable attendances** Identified using methodology developed by University of Sheffield's School of Health and Related Research's (ScHARR) definition of 'avoidable' Emergency Department attendances (Mason et al, 2017) with reference to the NHS Digital (2020) adaptation using SNOMED/ECDS coding.

-  **Divertible attendances** These patients are those who are are not avoidable attendances, minor acuity, minor investigations of PbR category 1 or 2, minor treatments of PbR category 1 or 2 and were discharged.

-  **Appropriate attendances** These who were discharged and did not meet the above 2 criteria.

-  **Admissions** These are patients who were admitted.

In addition to pure numbers of attendances we used a proxy for a measure of active or resource utilisation time for these patients.  So all time calculations take their total time in ward minus their wait time.  The study appreciates that this may not be completely active time with a clinician, but tries to account for the time they utilise that is not wait time.

Geo spacial analysis utilsed a travel time to department based on the patients home LSOA and utilised the average day car journey time.  The cohort was split to within 30 minutes travel time and over 30 minutes.  

## Comparison to pre covid levels

There are two parts to this and the graphs demonstrate the difference.  The first is simply the count by number of attendances.  This metric shows that broadly this metric of activity is in line with pre covid levels.

What is more diverse is the change of 'time in department'.  For this metric we took the total time in department minus the waiting time.  This was served as a proxy metric for time utilsing department resource.  The majority of sites have seen a sharper increase in this metric, which may be part of the reason that departments are reporting as busier.  This additional time may also correlate with more complex patient presentations, however this analysis does not explore that hypothesis.

We can also observe that this change is different across different presentation types.

## Comparison of increase of number of attendances and time (total - wait) {.tabset .tabset-fade .tabset-pills}

### Total type 1 walk in
```{r, echo=FALSE, fig.height=12,fig.width = 16}
grid.arrange(change_plot_ts(attend_pat_cln, 'Total', 'attends'),
             change_plot_ts(attend_pat_cln, 'Total', 'resource_time'),
             ncol = 2)
```

### `r presentaion_popular_site$chiefcomplaintgrouping[1]`
```{r, echo=FALSE, fig.height=12,fig.width = 16}
grid.arrange(change_plot_ts(attend_pat_cln, presentaion_popular_site$chiefcomplaintgrouping[1], 'attends'),
             change_plot_ts(attend_pat_cln, presentaion_popular_site$chiefcomplaintgrouping[1], 'resource_time'),
             ncol = 2)
```


## Benchmarking change over time {.tabset .tabset-fade .tabset-pills}

These charts look at the last 6 months of attendances and compare them to the same 6 months pre covid.  The percentage change is then calculated.  This is broken down across the various attendance types we have identified.  Each chart shows one of the top 6 chief complain presentation types.

We are most interested in this study for the change in divertible attendances.The other types are provided for context.  Further in the report we can see this in a more traditional time series for this individual site.

This broadly matches the previous charts, except the previous charts are accumulated the 6 month average whilst these charts are based on the aggregate for the 6 months. 

### Total Attendances
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_to


```

### Avoidable Attendances
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_aa


```

### Divertible Attendances
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_da


```


### Appropriate Attendances (Discharged)
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_ap 


```


### Appropriate Attendances (Admissions)
```{r, echo=FALSE, fig.height=12,fig.width = 16}
bench_ad 


```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

! *There are some data quality issues in some of the sites that may mean some of the cohorts may not be clearly identified.*   

</div>


## Overview by site
```{r, echo=FALSE, fig.height=12,fig.width = 16}
data_summary_table


```


## Resource time {.tabset .tabset-fade .tabset-pills}

### Resource time site comparison

The below chart shows the average resource time per patient for each
site over the last 18 months.

*'Resource time' is defined as total time in ward minus wait time.*

Guide for interpreting the box plot:

**Key Components of a Box Plot:**

**Box:** The central rectangle represents the interquartile range (IQR),
which contains the middle 50% of the data.

**Median:** The line within the box indicates the median value,
separating the lower and upper halves of the data.

**Whiskers:** The lines extending from the box are called whiskers. They
typically reach to the minimum and maximum values excluding outliers.

**Outliers:** Data points that fall outside of the whiskers, represented
by the dots, are considered outliers.

```{r echo=FALSE, results='hide', fig.height=6,fig.width = 12,message=FALSE, warnings = FALSE}
boxplot_allsites

```

### Resource time site level time comparison for 'divertible patients'

This shows the distribution of resource time over six month time periods. 

```{r echo=FALSE, results='hide',fig.height=6,fig.width = 12,message=FALSE, warnings = FALSE}

boxplot_month

```

### Resource time SPC for 'divertible patients'

The below SPC chart shows the median resource time per patient over the
latest 18 month for this site.

```{r echo=FALSE, results='hide', fig.height=6,fig.width = 12,message=FALSE, warnings = FALSE}
plot_SPC_Resource_Median

```

Statistical Process Control charts use statistics to identify patterns and anomalies that may be worth investigating as they are not caused by common cause variation within the data. 

## Chief Complaint

The below SPC chart shows the number of attendances by the top 6 
complaint times over the latest 18 month for this site.

```{r echo=FALSE, results='hide',fig.height=6, fig.width = 12,message=FALSE, warnings = FALSE}
CG_SPC

```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
*Please note that Y axis scales are different for each graph.*
</div>

## Clustering

We have explored the divertible attendances and by running correlation models across the features of the patients, we have identified the three strongest characteristics of these patients.  These are their age, the time that they attend A&E and their travel time to the hospital.

These tables give a baseline of the all the attendances for patients for the chief complaint, we then compare this to the characteristics of the divertible attendance group, further we explore if these features are further correlated nd have employed a clustering algorithm  to group the patients into similar cohorts.  This pulls together into groups where features are most strongly correlated.

### Breakdown by chief complaint  {.tabset .tabset-fade .tabset-pills}

#### `r chief_compaints_list[1]`
```{r, echo=FALSE, warning=FALSE, fig.height=12,fig.width = 16}
create_cluster_table(chief_compaints_list[1])

```

*Explanation of cluster groups.*  


`r comments(chief_compaints_list[1])[1]`  

`r comments(chief_compaints_list[1])[2]`  

`r comments(chief_compaints_list[1])[3]`  


#### `r chief_compaints_list[2]`
```{r, echo=FALSE, warning=FALSE, fig.height=12,fig.width = 16}
create_cluster_table(chief_compaints_list[2])

```

*Explanation of cluster groups.*  

`r comments(chief_compaints_list[2])[1]`  

`r comments(chief_compaints_list[2])[2]`  

`r comments(chief_compaints_list[2])[3]`  

#### `r chief_compaints_list[3]`
```{r, echo=FALSE, warning=FALSE, fig.height=12,fig.width = 16}
create_cluster_table(chief_compaints_list[3])

```

*Explanation of cluster groups.*  

`r comments(chief_compaints_list[3])[1]`  

`r comments(chief_compaints_list[3])[2]`  

`r comments(chief_compaints_list[3])[3]`  

#### `r chief_compaints_list[4]`
```{r, echo=FALSE, warning=FALSE, fig.height=12,fig.width = 16}
create_cluster_table(chief_compaints_list[4])

```

*Explanation of cluster groups.*  

`r comments(chief_compaints_list[4])[1]`  

`r comments(chief_compaints_list[4])[2]`  

`r comments(chief_compaints_list[4])[3]`  

#### `r chief_compaints_list[5]`
```{r, echo=FALSE, warning=FALSE, fig.height=12,fig.width = 16}
create_cluster_table(chief_compaints_list[5])

```

*Explanation of cluster groups.*  

`r comments(chief_compaints_list[5])[1]`  

`r comments(chief_compaints_list[5])[2]`  

`r comments(chief_compaints_list[5])[3]`  

#### `r chief_compaints_list[6]`
```{r, echo=FALSE, warning=FALSE, fig.height=12,fig.width = 16}
create_cluster_table(chief_compaints_list[6])

```

*Explanation of cluster groups.*  

`r comments(chief_compaints_list[1])[6]`  

`r comments(chief_compaints_list[1])[6]`  

`r comments(chief_compaints_list[1])[6]`  




## Appendix:

Data sources: [Reporting].[MESH_ECDS_EC_Core],[Reporting_UKHD_ODS].[Provider_Site], [AGEM_TravelTime].[LSOA_to_Site]

Time period: 18 months from lastest month end. 

Avoidable attendances is based on work by NHS Digital in 2020:  

<https://web.archive.org/web/20210226213236/https:/digital.nhs.uk/data-and-information/data-tools-and-services/data-services/innovative-uses-of-data/demand-on-healthcare/unnecessary-a-and-e-attendances>


Resource time: Total time to discharge - Time to treat. 95th percentile
was used to account for outliers. Any patients where their time to treat
was longer than their discharge time were removed.


## Contact

> [**Mark Crawford**](mailto:mark.crawford@nhs.net)\
> [**Chris Brownett**](mailto:chris.brownett@nhs.net)\
> [**Alistair Hewson**](mailto:alistair.hewson@nhs.net)\
> [**Tomasz Redka**](mailto:tomaz.redka@nhs.net)\
> [**Simon Wellesley-Miller**](mailto:Simon.Wellesley-Miller@nhs.net)\
> *South West Insights and Intelligence Team*\
> Version 0.9\ 
> 13 September 2024\ 
